PyQBench: a Python library for benchmarking gate-based quantum computers

### 2023-02-13 txt/manuscript.tex
add examples

%
%\lstdefinestyle{mystyle}{
%	backgroundcolor=\color{backcolour},
%	commentstyle=\color{codegreen},
%	keywordstyle=\color{magenta},
%	numberstyle=\tiny\color{codegray},
%	stringstyle=\color{codepurple},
%	basicstyle=\ttfamily\footnotesize,
%	breakatwhitespace=false,
%	breaklines=true,
%	captionpos=b,
%	keepspaces=true,
%	numbers=left,
%	numbersep=5pt,
%	showspaces=false,
%	showstringspaces=false,
%	showtabs=false,
%	tabsize=2
%}
%
%\lstset{style=mystyle}
\lstset{frame=tb,
	language=Python,
	aboveskip=3mm,
	belowskip=3mm,
	showstringspaces=false,
	columns=flexible,
	basicstyle={\small\ttfamily},
	numberstyle=\color{blue},
	keywordstyle=\color{blue},
	commentstyle=\color{PineGreen},
	stringstyle=\color{ForestGreen},
	breaklines=true,
	breakatwhitespace=true,
	tabsize=3,
	escapeinside={@}{@},
	literate=
	{0}{{{\color{red}0}}}1
	{1}{{{\color{red}1}}}1
	{2}{{{\color{red}2}}}1
	{3}{{{\color{red}3}}}1
	{4}{{{\color{red}4}}}1
	{5}{{{\color{red}5}}}1
	{6}{{{\color{red}6}}}1
	{7}{{{\color{red}7}}}1
	{8}{{{\color{red}8}}}1
	{9}{{{\color{red}9}}}1
    {v0}{{{\color{black}v0}}}1 %% we can not style only numeric literals
    {v1}{{{\color{black}v1}}}1
}
\restylefloat{table}

@@

Now we simulate  theoretical framework using PyQBench. We will start with simulation in Qiskit SDK with Aer simulator. First, we import necessary modules. \todo[inline]{czy to zdanie jest poprawne?} 
\todo[inline]{listingi swieca sie na razie jak choinka ale bedziemy kolorki drasowac}

@@

\begin{lstlisting}[language=Python, caption=Defining the experiment]
type: discrimination-fourier
qubits:
	- target: 0
	  ancilla: 1
	- target: 2
      ancilla: 3
angles:
	start: 0
	stop: 2 * pi
	num_steps: 50
gateset: ibmq
method: direct_sum
num_shots: 100
\end{lstlisting}

The components of the experiment is as follows:
\begin{itemize}
	\item The \textit{type} describes type of the experiment. Currently, the only option for \textit{type} is \textit{discrimination-fourier}. In the future, other benchmarks may be added, in which keys the \textit{types} key should contain their name.
	\begin{lstlisting}[language=Python]
type: discrimination-fourier
	\end{lstlisting}
	\item The \textit{qubits} list describes pair of qubits on which the experiment should be run. In our example, the benchmark will run on two pairs of qubits. First pair is 0 and 1, second pair is 2 and 3.
	Notice that we distinguish between ancilla and target qubit, so e.g

@@

Observe that the experiment file does not specify what backend to use. This way, the same experiment file can be used on multiple backends.
 Different backends may require different information to be used. For instance, IBMQ backends might require hub, group and project, whereas backends from the \textit{Aer} package do not. Hence, the exact format of the YAML file depends on what backend one wants to use. Below we present 
 YAML file defining the IMBQ backend named Quito. 
\begin{lstlisting}[language=Python, caption=IBMQ backend]
name: ibmq_quito
asynchronous: true
provider:
	hub: ibm-q
	group: open
	project: main
\end{lstlisting}

@@

\subsubsection{Running the experiment and collecting measurements data}
After preparing YAML files with defined experiment and backend, 
running the experiment is done by using the following command line invocation:
\begin{lstlisting}[language=Python]
qbench disc-fourier benchmark experiment_file.yml backend_file.yml
\end{lstlisting}
The output file will be printed to stdout. Optionally, the \textit{- -output OUTPUT} flag might be provided to write the output to the \textit{OUTPUT} file instead.
\begin{lstlisting}[language=Python]
qbench disc-fourier benchmark experiment_file.yml backend_file.yml --output async_results.yml
\end{lstlisting}
The result of running the above command can be twofold:
\begin{itemize}
	\item If backend is asynchronous, the output will contain intermediate data containing, amongst others, job$\_$ids correlated with the circuit they correspond to.
	\item If the backend is synchronous, the output will contain measurement data (bitstrings) for each of the circuits run.
\end{itemize}
For asynchronous experiments, the output is as follows:
\todo[inline]{result}

For synchronous experiment, the output looks similar to the one below: 
\todo[inline]{result}

@@

\subsubsection{Resolving asynchronous jobs}
If the status of jobs is \textit{DONE}, we have to obtain measurements from the submitted jobs. This is done using the following command:
\begin{lstlisting}[language=Python]
qbench disc-fourier resolve async-results.yml resolved.yml
\end{lstlisting}
The resolved results, stored in \textit{resolved.yml}, would look just like the experiment was run synchronously. In other words, we should have a file containing measurement histograms, no matter if we run a synchronous or asynchronous experiment. After this step, we are ready to computing probabilities.

\subsubsection{Computing probabilities}
To construct a table with discrimination probabilities, you can use the following command
\begin{lstlisting}[language=Python]
qbench disc-fourier tabulate resolved.yml results.csv 
\end{lstlisting}
The results are saved as file csv as a table.
A single row for each tuple of \textit{(target, ancilla, phi)}. The \textit{disc\_prob} column will contain the discrimination probability for given configuration. The \textit{mit\_disc\_prob} column contains discrimination probability if Mthree mitigation was applied~\cite{mthree, mthreepublication}.

@@

>>>>

%
%\lstdefinestyle{mystyle}{
%	backgroundcolor=\color{backcolour},
%	commentstyle=\color{codegreen},
%	keywordstyle=\color{magenta},
%	numberstyle=\tiny\color{codegray},
%	stringstyle=\color{codepurple},
%	basicstyle=\ttfamily\footnotesize,
%	breakatwhitespace=false,
%	breaklines=true,
%	captionpos=b,
%	keepspaces=true,
%	numbers=left,
%	numbersep=5pt,
%	showspaces=false,
%	showstringspaces=false,
%	showtabs=false,
%	tabsize=2
%}
%
%\lstset{style=mystyle}
\lstset{frame=tb,
	language=Python,
	aboveskip=3mm,
	belowskip=3mm,
	showstringspaces=false,
	columns=flexible,
	basicstyle={\small\ttfamily},
%	numberstyle=\color{blue},
%	keywordstyle=\color{blue},
%	commentstyle=\color{PineGreen},
%	stringstyle=\color{ForestGreen},
	breaklines=true,
	breakatwhitespace=true,
	tabsize=3,
	escapeinside={@}{@},
%	literate=
%	{0}{{{\color{red}0}}}1
%	{1}{{{\color{red}1}}}1
%	{2}{{{\color{red}2}}}1
%	{3}{{{\color{red}3}}}1
%	{4}{{{\color{red}4}}}1
%	{5}{{{\color{red}5}}}1
%	{6}{{{\color{red}6}}}1
%	{7}{{{\color{red}7}}}1
%	{8}{{{\color{red}8}}}1
%	{9}{{{\color{red}9}}}1
%    {v0}{{{\color{black}v0}}}1 %% we can not style only numeric literals
%    {v1}{{{\color{black}v1}}}1
}
\restylefloat{table}

@@

Now we simulate  theoretical framework using PyQBench. We will start with simulation in Qiskit SDK with Aer simulator. First, we import necessary modules. \todo[inline]{czy to zdanie jest poprawne?}

@@

\begin{lstlisting}[language=Python, caption=Defining the experiment]
type: discrimination-fourier
qubits:
	- target: 0
	  ancilla: 1
	- target: 1
      ancilla: 2
angles:
	start: 0
	stop: 2 * pi
	num_steps: 3
gateset: ibmq
method: direct_sum
num_shots: 100
\end{lstlisting}

The components of the experiment is as follows:
\begin{itemize}
	\item The \textit{type} describes type of the experiment. Currently, the only option for \textit{type} is \textit{discrimination-fourier}. In the future, other benchmarks may be added, in which keys the \textit{types} key should contain their name.
	\begin{lstlisting}[language=Python]
type: discrimination-fourier
	\end{lstlisting}
	\item The \textit{qubits} list describes pair of qubits on which the experiment should be run. In our example, the benchmark will run on two pairs of qubits. First pair is 0 and 1, second pair is 1 and 2.
	Notice that we distinguish between ancilla and target qubit, so e.g

@@

Observe that the experiment file does not specify what backend to use. This way, the same experiment file can be used on multiple backends.
 Different backends may require different information to be used. For instance, IBMQ backends might require hub, group and project, whereas backends from the \textit{Aer} package do not. Hence, the exact format of the YAML file depends on what backend one wants to use. Below we present 
 YAML file defining the IMBQ backend named Quito. 
\begin{lstlisting}[language=Python, caption=IBMQ backend]
name: ibmq_quito
asynchronous: false
provider:
	hub: ibm-q
	group: open
	project: main
\end{lstlisting}

@@

\subsubsection{Running the experiment and collecting measurements data}
After preparing YAML files with defined experiment and backend, 
running the experiment is done by using the following command line invocation:
\begin{lstlisting}[language=Python]
qbench disc-fourier benchmark experiment_file.yml backend_file.yml
\end{lstlisting}
The output file will be printed to stdout. Optionally, the \textit{- -output OUTPUT} flag might be provided to write the output to the \textit{OUTPUT} file instead.
\begin{lstlisting}[language=Python]
qbench disc-fourier benchmark experiment_file.yml backend_file.yml --output async_results.yml
\end{lstlisting}
The result of running the above command can be twofold:
\begin{itemize}
	\item If backend is asynchronous, the output will contain intermediate data containing, amongst others, job$\_$ids correlated with the circuit they correspond to.
	\item If the backend is synchronous, the output will contain measurement data (bitstrings) for each of the circuits run.
\end{itemize}

For synchronous experiment, the output looks similar to the one below: 
\begin{lstlisting}[language=Python]
metadata:
	experiments:
		type: discrimination-fourier
		qubits:
		- {target: 0, ancilla: 1}
		- {target: 1, ancilla: 2}
		angles: {start: 0.0, stop: 6.283185307179586, num_steps: 3}
		gateset: ibmq
		method: direct_sum
		num_shots: 100
	backend_description:
		name: ibmq_quito
		asynchronous: false
		provider: {group: open, hub: ibm-q, project: main}
data:
- target: 0
  ancilla: 1
  phi: 0.0
  results_per_circuit:
  - name: id
  histogram: {'00': 28, '01': 26, '10': 21, '11': 25}
  mitigation_info:
	target: {prob_meas0_prep1: 0.052200000000000024, prob_meas1_prep0: 0.0172}
	ancilla: {prob_meas0_prep1: 0.05900000000000005, prob_meas1_prep0: 0.0202}
  mitigated_histogram: {'00': 0.2637212373658018, '01': 0.25865061319892463, '10': 0.2067279352110304, '11': 0.2709002142242433}
  - name: u
  histogram: {'00': 30, '01': 16, '10': 28, '11': 26}
  mitigation_info:
	target: {prob_meas0_prep1: 0.052200000000000024, prob_meas1_prep0: 0.0172}
	ancilla: {prob_meas0_prep1: 0.05900000000000005, prob_meas1_prep0: 0.0202}
  mitigated_histogram: {'00': 0.2857378991684036, '01': 0.14975297832942433, '10': 0.28142307224788693, '11': 0.2830860502542851}
- target: 0
  ancilla: 1
  phi: 3.141592653589793
  results_per_circuit:
  - name: id
  histogram: {'00': 4, '01': 5, '10': 45, '11': 46}
  mitigation_info:
	target: {prob_meas0_prep1: 0.052200000000000024, prob_meas1_prep0: 0.0172}
	ancilla: {prob_meas0_prep1: 0.05900000000000005, prob_meas1_prep0: 0.0202}
  mitigated_histogram: {'00': 0.011053610583159325, '01': 0.02261276648026373, '10': 0.4593955619936729, '11': 0.5069380609429042}
  - name: u
  histogram: {'00': 56, '01': 43, '10': 1}
  mitigation_info:
  target: {prob_meas0_prep1: 0.052200000000000024, prob_meas1_prep0: 0.0172}
  ancilla: {prob_meas0_prep1: 0.05900000000000005, prob_meas1_prep0: 0.0202}
  mitigated_histogram: {'00': 0.5573987337172156, '01': 0.44424718645642625, '10': -0.0016459201736417181}
  - target: 0
    ancilla: 1
  phi: 6.283185307179586
  results_per_circuit:
  - name: id
  histogram: {'00': 36, '01': 18, '10': 25, '11': 21}
  mitigation_info:
	target: {prob_meas0_prep1: 0.052200000000000024, prob_meas1_prep0: 0.0172}
	ancilla: {prob_meas0_prep1: 0.05900000000000005, prob_meas1_prep0: 0.0202}
  mitigated_histogram: {'00': 0.3488190312089973, '01': 0.17355281935572894, '10': 0.2505792064871127, '11': 0.22704894294816103}
  - name: u
  histogram: {'00': 32, '01': 27, '10': 24, '11': 17}
  mitigation_info:
	target: {prob_meas0_prep1: 0.052200000000000024, prob_meas1_prep0: 0.0172}
	ancilla: {prob_meas0_prep1: 0.05900000000000005, prob_meas1_prep0: 0.0202}
  mitigated_histogram: {'00': 0.3025357275361897, '01': 0.27413673119534815, '10': 0.24313373302688793, '11': 0.18019380824157433}
- target: 1
  ancilla: 2
  phi: 0.0
  results_per_circuit:
  - name: id
  histogram: {'00': 27, '01': 20, '10': 24, '11': 29}
  mitigation_info:
	target: {prob_meas0_prep1: 0.05900000000000005, prob_meas1_prep0: 0.0202}
	ancilla: {prob_meas0_prep1: 0.07540000000000002, prob_meas1_prep0: 0.0528}
  mitigated_histogram: {'00': 0.2594378169217188, '01': 0.19318893233269735, '10': 0.23035366874292057, '11': 0.3170195820026633}
  - name: u
  histogram: {'00': 31, '01': 24, '10': 23, '11': 22}
  mitigation_info:
	target: {prob_meas0_prep1: 0.05900000000000005, prob_meas1_prep0: 0.0202}
	ancilla: {prob_meas0_prep1: 0.07540000000000002, prob_meas1_prep0: 0.0528}
  mitigated_histogram: {'00': 0.30056875246775644, '01': 0.2438221628798003, '10': 0.22180309809696985, '11': 0.23380598655547338}
- target: 1
  ancilla: 2
  phi: 3.141592653589793
  results_per_circuit:
  - name: id
  histogram: {'00': 5, '01': 4, '10': 50, '11': 41}
  mitigation_info:
	target: {prob_meas0_prep1: 0.05900000000000005, prob_meas1_prep0: 0.0202}
	ancilla: {prob_meas0_prep1: 0.07540000000000002, prob_meas1_prep0: 0.0528}
  mitigated_histogram: {'00': 0.009552870928837118, '01': 0.007194089383161034,
  '10': 0.5236791012692514, '11': 0.4595739384187503}
  - name: u
  histogram: {'00': 41, '01': 51, '10': 3, '11': 5}
  mitigation_info:
	target: {prob_meas0_prep1: 0.05900000000000005, prob_meas1_prep0: 0.0202}
	ancilla: {prob_meas0_prep1: 0.07540000000000002, prob_meas1_prep0: 0.0528}
  mitigated_histogram: {'00': 0.4073387714165384, '01': 0.5614614121117936, '10': 0.006431862814564833, '11': 0.024767953657102992}
- target: 1
  ancilla: 2
  phi: 6.283185307179586
  results_per_circuit:
  - name: id
  histogram: {'00': 30, '01': 28, '10': 23, '11': 19}
  mitigation_info:
	target: {prob_meas0_prep1: 0.05900000000000005, prob_meas1_prep0: 0.0202}
	ancilla: {prob_meas0_prep1: 0.07540000000000002, prob_meas1_prep0: 0.0528}
  mitigated_histogram: {'00': 0.2868459834940102, '01': 0.2919564941384742, '10': 0.22466574543735374, '11': 0.19653177693016174}
  - name: u
  histogram: {'00': 15, '01': 20, '10': 36, '11': 29}
  mitigation_info:
	target: {prob_meas0_prep1: 0.05900000000000005, prob_meas1_prep0: 0.0202}
	ancilla: {prob_meas0_prep1: 0.07540000000000002, prob_meas1_prep0: 0.0528}
  mitigated_histogram: {'00': 0.1187719606657805, '01': 0.1962085394489247, '10': 0.3710195249988589, '11': 0.31399997488643583}
Consider the particular experiment data
\begin{lstlisting}[language=Python]
data:
- target: 0
  ancilla: 1
  phi: 0.0
  results_per_circuit:
  - name: id
  histogram: {'00': 28, '01': 26, '10': 21, '11': 25}
  mitigation_info:
	target: {prob_meas0_prep1: 0.052200000000000024, prob_meas1_prep0: 0.0172}
	ancilla: {prob_meas0_prep1: 0.05900000000000005, prob_meas1_prep0: 0.0202}
  mitigated_histogram: {'00': 0.2637212373658018, '01': 0.25865061319892463, '10': 0.2067279352110304, '11': 0.2709002142242433}
	The data includes \textit{target, ancilla, phi},
	and \textit{results\_pre\_circuit}.   The first three pieces of information have already been described. The last data  \textit{results\_pre\_circuit} gives us 
	the following additional informations: 
	\begin{itemize}
		\item  name: id:  the information which measurement is used during experiment,   $\PP_{U}$ or $\PP_{\Id}$. In this example we consider  $\PP_{\Id}$. 
		\item histogram: {'00': 28, '01': 26, '10': 21, '11': 25}: the dictionary with measurement's outcomes. The keys represent possible measurement oucomes, whereas the values present number of occurrences. 
		\item mitigation\_info:  ??????????  ~\cite{mthree, mthreepublication}
		\item mitigated\_histogram:  the histogram with measurement's outcomes after mitigation.
	\end{itemize}

@@

\subsubsection{Resolving asynchronous jobs}
If the status of jobs is \textit{DONE}, we have to obtain measurements from the submitted jobs. This is done using the following command:
\begin{lstlisting}[language=Python]
qbench disc-fourier resolve async-results.yml resolved.yml
\end{lstlisting}

The resolved results, stored in \textit{resolved.yml}, would look just like the experiment was run synchronously. In other words, we should have a file containing measurement histograms, no matter if we run a synchronous or asynchronous experiment (see \ref{app:example}). Now, we are ready to computing probabilities.

\subsubsection{Computing probabilities}
Regardless of whether we use a synchronous or asynchronous approach, we can convert the resulting YML file \textit{results.yml} to a CSV file by creating a result table in which we place the results of discrimination probabilities.
To construct a table with probabilities, you can use the following command
\begin{lstlisting}[language=Python]
qbench disc-fourier tabulate results.yml results.csv 
\end{lstlisting}
A single row for each tuple of \textit{(target, ancilla, phi)}. The \textit{disc\_prob} column will contain the discrimination probability for given configuration. The \textit{mit\_disc\_prob} column contains discrimination probability if Mthree mitigation was applied~\cite{mthree, mthreepublication}.
A sample csv file is provided below:
\begin{figure}[h!]
	\centering
	\includegraphics[scale=1]{results/async_result}
	\caption{Tabulate results}
	\label{fig:tabulateresults}
\end{figure}

@@

\section{Example of files} \label{app:example}
